Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-03 Multiresolution_framework/ltx/manuscript.tex
changed the figure for the kernel estimate and deleted extra figs
changed field estimation plot and deleted extra figures
changed figure name in tex file
changed all the equations to Eq.

\section{Multi-Resolution Cortical Architecture}
The human neocortex has an inherent multi-resolution architecture. The spatiotemporal neural field can be written as 
\begin{equation}
	v\left(\mathbf{r},t\right) = \sum_{j} v_j\left(\mathbf{r},t\right),
\end{equation}
where $v\left(\mathbf{r},t\right)$ is a voltage field that may be measured with a multi-electrode array or voltage sensitive dye, $\mathbf{r}\in \mathbb{R}^n$ (where $n\le3$) describes the spatial location within the field, $t$ denotes time, and $j$ indexes the different levels of resolution in the field. This multi-resolution structure is depicted in Fig.~\ref{fig:MultiLayerFieldModel}.
\begin{figure}[t]
	\centering
		\includegraphics[scale=1]{./Graph/MultiResNeuralField.pdf}
	\caption{{\bf Multi-layer neural field model}. The inset figures show the shape of the
connectivity kernel components at each spatial resolution.}
	\label{fig:MultiLayerFieldModel}
\end{figure}

@@

The stochastic IDE form of the Amari neural field formulation~\cite{Amari1977} is given by (see~\cite{Freestone2011} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, \mathrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where in this first-order (in time) form of the model the membrane dynamics are included through the parameter $\xi=1-T_s/\tau$, where $\tau$ is the membrane time constant and $T_s$ is the sampling time. 
% The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the connectivity kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977,Atay2005}. The components of the model are shown in Fig.~\ref{fig:Modelcomponents}
\begin{figure}[!t]
\centering
\includegraphics[scale=1]{./Graph/ModelComponents.pdf}
\caption{ {\bf The IDE neural field model components}. (\textbf{A}) The Mexican hat (dotted line) and Wizard hat (solid line) connectivity kernels. (\textbf{B}) First order (dotted line) and second order (solid line) synaptic kernels. (\textbf{C}) A representative of the activation function.}
\label{fig:Modelcomponents}
\end{figure}

@@

Now we have the compact form of the decomposition we can substitute this into the neural field equation \eqref{eq:DiscreteTimeModel} giving

@@

The aim of the next part of the derivation is to isolate the state vector on the left hand side of the equation. To do this we first multiply both sides of equation~\eqref{eq:ApproxDiscreteTimeModel} by $\boldsymbol\mu\left(r\right)$ and integrate over space giving 

@@

and substitute into equation~\eqref{eq:ApproxDiscreteTimeModel2} giving

@@

\parham{The state-transition equation can be simplified if the chosen family of wavelets are isotropic where}
\begin{equation}
	\boldsymbol{\lambda}_{j,l}(r-r') = \boldsymbol{\lambda}(2\mathbf{c}_{j,l}+r'-r). 
\end{equation}
Following this, we can switch the order of operations with the inner product and convolution in equation~\eqref{eq:ApproxDiscreteTimeModel4}. With this switch the convolution is now between the scaling functions and wavelets for the field and the scaling functions and wavelets for the connectivity kernel. This convolution can be calculated analytically when using B-spline wavelets and can be predefined once, since these are static functions. This has major computation advantages when solving the estimation problem. \parham{Although this switch holds theoretically it should be noted that it will introduce numerical error, specially close to the edges, due to the finite extent of the field. } The predefined convolution can be written as
\begin{equation}\label{eq:Gammaij}
	\left[\Gamma(r')\right]_{:j,l} = T_s \mathbf{\Lambda}_{x}^{-1}\int_{\Omega} \boldsymbol\mu\left(r\right)\boldsymbol\lambda\left(2c_{j,l} + r'-r\right) \,\mathrm{d}r,
\end{equation}
giving
\begin{equation}
	\mathbf{x}_{t+1} = 
	\xi \mathbf{x}_t + 
	\int_{\Omega} \Gamma\left(r'\right)f\left(\boldsymbol\mu^\top\left(r'\right) \mathbf{x}_t\right) 
	\, \mathrm{d}r' \boldsymbol\theta
	+ \mathbf w_t.
\end{equation}

% Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
% \begin{align}\label{eq:DecomposedModel2} 
% 	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1} = 
% 	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
% \end{align}
% Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
% \begin{align}\label{eq:StateEquation}
%  \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
% \label{eq:A_theta}
%  \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
% \end{align}
% The disturbance becomes 
% \begin{equation}\label{eq:Disturbance}
% \mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
% \end{equation}
% which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
% \begin{equation}\label{eq:CovMatrix}
% \boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
% \end{equation}

The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix, $\mathbf{C}$, is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ik} \triangleq \int_{\Omega}m(r_i -r')\boldsymbol{\mu}_k(r') \, \mathrm{d}r'.
\end{equation}

\section{B-Spline MRAIDE in State-Space }
\subsection{A Brief Introduction to B-Spline Wavelets}
B-spline functions are one of the most basic elements for wavelet construction. These functions are composed of piecewise polynomials combined smoothly at some breaking points (knots), which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \citep{Goswami1999}, where the higher the order the smoother the resulting B-spline becomes (see Fig.~\ref{fig:Figure0}).
\begin{figure}[!t]
\centering
\includegraphics{./Graph/Figure0.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (colored lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}

@@

The convolution and inner product of B-spline functions are required in order to formulate the multi-resolution framework described in the previous section (construction of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Gamma$). To show how the convolution of two B-splines is calculated, \eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself
\begin{equation}\label{eq:N1convolutions}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of \eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e. $ N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right)$ (see Appendix~\ref{ap:InnerProductOfBsplines} for explicit derivation). The values of $N_{m+m'}$ in \eqref{eq:BsplineInnerProduct} can be easily determined recursively by evaluating \eqref{eq:MRA-DoBoorFormula} at integer points, i.e.
 \begin{equation}\label{eq:MRA-recursiveBsplineatintegerpoints}
 \begin{cases}
 N_2(k)=\delta(k-1)\quad k\in \mathbb{Z}, \\
 N_{m}\left(k\right)=\frac{k}{m-1}N_{m-1}\left(k\right)+\frac{m-k}{m-1}N_{m-1}\left(k-1\right) \quad k=1,2,\dots,m-1
  \end{cases}
 \end{equation}
Note $N_{m}\left(k\right)=0$ for $k\le0$ or $k\ge m$.

@@

Fig.~\ref{fig:MRA-Figure1} shows the cubic B-spline scaling function and its associated wavelet function with dilation and translation parameters set to zero. Note in general the supports of the scaling and wavelet functions are given by 
\begin{align}
	\mathrm{supp}(N_{m;j,l})&=\left[\frac{l}{2^j},\frac{m+l}{2^j}\right] \\  
  \mathrm{supp}(\varphi_{m;j,l})&=\left[\frac{l}{2^j},\frac{2m-1+l}{2^j}\right] 
	\end{align} 
\begin{figure}[!t]
\centering
\includegraphics{./Graph/Figure1.pdf}
% \includegraphics{./Graph/Figure1b.pdf}
\caption{{\bf Examples of B-spline scaling and wavelet functions}. (\textbf{A}) The cubic B-spline scaling function. (\textbf{B}) The corresponding wavelet function.}
\label{fig:MRA-Figure1}
\end{figure}

By exploiting \eqref{eq:BsplineConvolution} and \eqref{eq:BsplineInnerProduct}, the integrals in \eqref{eq:Lambdax} and \eqref{eq:Gammaij} can be computed analytically. In constructing $\boldsymbol\Lambda_{x}$, it is important to note that B-spline scaling and wavelet functions possess the following orthogonality properties \citep{Unser1993}: 
\begin{equation}
  \left\langle \varphi_{m;j_1,l_1}(r),\varphi_{m;j_2,l_2}(r)\right\rangle =0  \quad \mathrm{for} \quad j_1\neq j_2
 \label{eq:MRA-PsiPsiOrthogonality} 
 \end{equation}
 \begin{equation}
  \left\langle N_{m;j_1,l_1}(r),\varphi_{m;j_2,l_2}(r)\right\rangle =0  \quad \mathrm{for} \quad j_1\leq j_2.
 \label{eq:MRA-PhiPsiOrthogonality}
 \end{equation}
  \parham{Note from \eqref{eq:MRA-PsiPsiOrthogonality} B-spline wavelets are semi-orthogonal as they are not orthogonal with respect to translation on a given scale.}

In order to analytically calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale using the two-scale relation pair defined in equations (\ref{eq:MRA-TwoScalepair1}-\ref{eq:MRA-TwoScalepair2coefs}). For scaling functions this can be done  by using \eqref{eq:MRA-TwoScalepair1} recursively, giving
\begin{align}\label{eq:MRA-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_j=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_j}N_m(2^jr-l_{n_1,n_2, \dots, n_j}),
\end{align}
where 
\begin{align}
 l_{n_1,n_2, \dots, n_j}=2^{j-1}n_1+2^{j-2}n_2+ \dots +2^{0}n_j.
\end{align}
The relation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} can be also proven by induction (see Appendix~\ref{ap:Induction}). A similar method can be used for wavelet functions noting that they must be written first in terms of scaling functions using \eqref{eq:MRA-TwoScalepair2}. Examples of such expansions are given in Fig.~\ref{fig:MRA-BasisDecomposition} where the scaling and wavelet functions at $j=0$ are written in terms of 625 and 1250 scaling functions at $j=4$ respectively.
\begin{figure}[!t] 
 	\centering
 		\includegraphics[scale=1]{./Graph/Decomposition.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. (\textbf{A}) The scaling function (blue) at $j=0$ is obtained using 625 scaling functions (black) at $j=4$. (\textbf{B}) The wavelet function (blue) at $j=0$ is obtained using 1250 scaling functions (black) at $j=4$.}
 	\label{fig:MRA-BasisDecomposition}
 \end{figure}  
 In this paper, 4th -- order cardinal B-spline scaling and wavelet functions are used, resulting in  8th and 12th order B-spline functions when calculating \eqref{eq:Lambdax} and \eqref{eq:Uij} respectively  whose values at integer points are given in Table~\ref{table:MRA-BsplineatIntegerPoints}.
\begin {table}[t]
\begin{center}
	\begin{tabular}{lcccc}
	\hline \hline
	& $k$ & $3!N_{4}\left(k\right)$ & $7!N_{8}\left(k\right)$ & $11!N_{12}\left(k\right)$\\ 
	\hline 
	& 1 & $1$ & $1$ & $1$\\
	& 2 & $4$ & $120$ & $2,036$\\
	& 3 & $\space$ & $1,191$ & $152,637$\\
	& 4 & $\space$ & $2,416$ & $2,203,488$\\
	& 5 & $\space$ & $\space$ & $9,738,114$\\
	& 6 & $\space$ & $\space$ & $15,724,248$\\
	\hline \hline
	\end{tabular}
 \caption {{\bf Cardinal B-Splines at the knot sequence}. Table entries are calculated using \eqref{eq:MRA-recursiveBsplineatintegerpoints} and can be also found in \citet{Goswami1999}.} 
 \label{table:MRA-BsplineatIntegerPoints}
 \end{center}
 \end {table}

>>>>

\section{Multi-Resolution Cortical Architecture}
The human neocortex has an inherent multi-resolution architecture. The spatiotemporal neural field can be written as 
\begin{equation}
	v\left(\mathbf{r},t\right) = \sum_{j} v_j\left(\mathbf{r},t\right),
\end{equation}
where $v\left(\mathbf{r},t\right)$ is a voltage field that may be measured with a multi-electrode array or voltage sensitive dye, $\mathbf{r}\in \mathbb{R}^n$ (where $n\le3$) describes the spatial location within the field, $t$ denotes time, and $j$ indexes the different levels of resolution in the field. This multi-resolution structure is depicted in Fig.~\ref{fig:MultiLayerFieldModel}.
\begin{figure}[t]
	\centering
		\includegraphics[scale=1]{./Graph/fig1.pdf}
	\caption{{\bf Multi-layer neural field model}. The inset figures show the shape of the
connectivity kernel components at each spatial resolution.}
	\label{fig:MultiLayerFieldModel}
\end{figure}

@@

The stochastic IDE form of the Amari neural field formulation~\cite{Amari1977} is given by (see~\cite{Freestone2011} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, \mathrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where in this first-order (in time) form of the model the membrane dynamics are included through the parameter $\xi=1-T_s/\tau$, where $\tau$ is the membrane time constant and $T_s$ is the sampling time. 
% The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the connectivity kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977,Atay2005}. The components of the model are shown in Fig.~\ref{fig:Modelcomponents}
\begin{figure}[!t]
\centering
\includegraphics[scale=1]{./Graph/fig2.pdf}
\caption{ {\bf The IDE neural field model components}. (\textbf{A}) The Mexican hat (dotted line) and Wizard hat (solid line) connectivity kernels. (\textbf{B}) First order (dotted line) and second order (solid line) synaptic kernels. (\textbf{C}) A representative of the activation function.}
\label{fig:Modelcomponents}
\end{figure}

@@

Now we have the compact form of the decomposition we can substitute this into the neural field Eq.~\eqref{eq:DiscreteTimeModel} giving

@@

The aim of the next part of the derivation is to isolate the state vector on the left hand side of the equation. To do this we first multiply both sides of Eq.~\eqref{eq:ApproxDiscreteTimeModel} by $\boldsymbol\mu\left(r\right)$ and integrate over space giving 

@@

and substitute into Eq.~\eqref{eq:ApproxDiscreteTimeModel2} giving

@@

\parham{The state-transition equation can be simplified if the chosen family of wavelets are isotropic where}
\begin{equation}
	\boldsymbol{\lambda}_{j,l}(r-r') = \boldsymbol{\lambda}(2\mathbf{c}_{j,l}+r'-r). 
\end{equation}
Following this, we can switch the order of operations with the inner product and convolution in Eq.~\eqref{eq:ApproxDiscreteTimeModel4}. With this switch the convolution is now between the scaling functions and wavelets for the field and the scaling functions and wavelets for the connectivity kernel. This convolution can be calculated analytically when using B-spline wavelets and can be predefined once, since these are static functions. This has major computation advantages when solving the estimation problem. \parham{Although this switch holds theoretically it should be noted that it will introduce numerical error, specially close to the edges, due to the finite extent of the field. } The predefined convolution can be written as
\begin{equation}\label{eq:Gammaij}
	\left[\Gamma(r')\right]_{:j,l} = T_s \mathbf{\Lambda}_{x}^{-1}\int_{\Omega} \boldsymbol\mu\left(r\right)\boldsymbol\lambda\left(2c_{j,l} + r'-r\right) \,\mathrm{d}r,
\end{equation}
giving
\begin{equation}
	\mathbf{x}_{t+1} = 
	\xi \mathbf{x}_t + 
	\int_{\Omega} \Gamma\left(r'\right)f\left(\boldsymbol\mu^\top\left(r'\right) \mathbf{x}_t\right) 
	\, \mathrm{d}r' \boldsymbol\theta
	+ \mathbf w_t.
\end{equation}

% Substituting Eq.~\eqref{eq:KernelFiniteExpansion} and Eq.~\eqref{eq:FieldFiniteExpansion} in Eq.~\eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
% \begin{align}\label{eq:DecomposedModel2} 
% 	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1} = 
% 	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
% \end{align}
% Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
% \begin{align}\label{eq:StateEquation}
%  \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
% \label{eq:A_theta}
%  \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
% \end{align}
% The disturbance becomes 
% \begin{equation}\label{eq:Disturbance}
% \mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
% \end{equation}
% which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
% \begin{equation}\label{eq:CovMatrix}
% \boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
% \end{equation}

The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into Eq.~\eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix, $\mathbf{C}$, is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ik} \triangleq \int_{\Omega}m(r_i -r')\boldsymbol{\mu}_k(r') \, \mathrm{d}r'.
\end{equation}

\section{B-Spline MRAIDE in State-Space }
\subsection{A Brief Introduction to B-Spline Wavelets}
B-spline functions are one of the most basic elements for wavelet construction. These functions are composed of piecewise polynomials combined smoothly at some breaking points (knots), which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \citep{Goswami1999}, where the higher the order the smoother the resulting B-spline becomes (see Fig.~\ref{fig:Figure0}).
\begin{figure}[!t]
\centering
\includegraphics{./Graph/fig3.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (colored lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}

@@

The convolution and inner product of B-spline functions are required in order to formulate the multi-resolution framework described in the previous section (construction of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Gamma$). To show how the convolution of two B-splines is calculated, Eq.~\eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself
\begin{equation}\label{eq:N1convolutions}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e. $ N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right)$ (see Appendix~\ref{ap:InnerProductOfBsplines} for explicit derivation). The values of $N_{m+m'}$ in Eq.~\eqref{eq:BsplineInnerProduct} can be easily determined recursively by evaluating Eq.~\eqref{eq:MRA-DoBoorFormula} at integer points, i.e.
 \begin{equation}\label{eq:MRA-recursiveBsplineatintegerpoints}
 \begin{cases}
 N_2(k)=\delta(k-1)\quad k\in \mathbb{Z}, \\
 N_{m}\left(k\right)=\frac{k}{m-1}N_{m-1}\left(k\right)+\frac{m-k}{m-1}N_{m-1}\left(k-1\right) \quad k=1,2,\dots,m-1
  \end{cases}
 \end{equation}
Note $N_{m}\left(k\right)=0$ for $k\le0$ or $k\ge m$.

@@

Fig.~\ref{fig:MRA-Figure1} shows the cubic B-spline scaling function and its associated wavelet function with dilation and translation parameters set to zero. Note in general the supports of the scaling and wavelet functions are given by 
\begin{align}
	\mathrm{supp}(N_{m;j,l})&=\left[\frac{l}{2^j},\frac{m+l}{2^j}\right] \\  
  \mathrm{supp}(\varphi_{m;j,l})&=\left[\frac{l}{2^j},\frac{2m-1+l}{2^j}\right] 
	\end{align} 
\begin{figure}[!t]
\centering
\includegraphics{./Graph/fig4.pdf}
% \includegraphics{./Graph/Figure1b.pdf}
\caption{{\bf Examples of B-spline scaling and wavelet functions}. (\textbf{A}) The cubic B-spline scaling function. (\textbf{B}) The corresponding wavelet function.}
\label{fig:MRA-Figure1}
\end{figure}

By exploiting Eq.~\eqref{eq:BsplineConvolution} and Eq.~\eqref{eq:BsplineInnerProduct}, the integrals in Eq.~\eqref{eq:Lambdax} and Eq.~\eqref{eq:Gammaij} can be computed analytically. In constructing $\boldsymbol\Lambda_{x}$, it is important to note that B-spline scaling and wavelet functions possess the following orthogonality properties \citep{Unser1993}: 
\begin{equation}
  \left\langle \varphi_{m;j_1,l_1}(r),\varphi_{m;j_2,l_2}(r)\right\rangle =0  \quad \mathrm{for} \quad j_1\neq j_2
 \label{eq:MRA-PsiPsiOrthogonality} 
 \end{equation}
 \begin{equation}
  \left\langle N_{m;j_1,l_1}(r),\varphi_{m;j_2,l_2}(r)\right\rangle =0  \quad \mathrm{for} \quad j_1\leq j_2.
 \label{eq:MRA-PhiPsiOrthogonality}
 \end{equation}
  \parham{Note from Eq.~\eqref{eq:MRA-PsiPsiOrthogonality} B-spline wavelets are semi-orthogonal as they are not orthogonal with respect to translation on a given scale.}

In order to analytically calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale using the two-scale relation pair defined in equations (\ref{eq:MRA-TwoScalepair1}-\ref{eq:MRA-TwoScalepair2coefs}). For scaling functions this can be done  by usingEq.~\eqref{eq:MRA-TwoScalepair1} recursively, giving
\begin{align}\label{eq:MRA-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_j=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_j}N_m(2^jr-l_{n_1,n_2, \dots, n_j}),
\end{align}
where 
\begin{align}
 l_{n_1,n_2, \dots, n_j}=2^{j-1}n_1+2^{j-2}n_2+ \dots +2^{0}n_j.
\end{align}
The relation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} can be also proven by induction (see Appendix~\ref{ap:Induction}). A similar method can be used for wavelet functions noting that they must be written first in terms of scaling functions using Eq.~\eqref{eq:MRA-TwoScalepair2}. Examples of such expansions are given in Fig.~\ref{fig:MRA-BasisDecomposition} where the scaling and wavelet functions at $j=0$ are written in terms of 625 and 1250 scaling functions at $j=4$ respectively.
\begin{figure}[!t] 
 	\centering
 		\includegraphics[scale=1]{./Graph/fig5.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. (\textbf{A}) The scaling function (blue) at $j=0$ is obtained using 625 scaling functions (black) at $j=4$. (\textbf{B}) The wavelet function (blue) at $j=0$ is obtained using 1250 scaling functions (black) at $j=4$.}
 	\label{fig:MRA-BasisDecomposition}
 \end{figure}  
 In this paper, 4th -- order cardinal B-spline scaling and wavelet functions are used, resulting in  8th and 12th order B-spline functions when calculating Eq.~\eqref{eq:Lambdax} and Eq.~\eqref{eq:Uij} respectively  whose values at integer points are given in Table~\ref{table:MRA-BsplineatIntegerPoints}.
\begin {table}[t]
\begin{center}
	\begin{tabular}{lcccc}
	\hline \hline
	& $k$ & $3!N_{4}\left(k\right)$ & $7!N_{8}\left(k\right)$ & $11!N_{12}\left(k\right)$\\ 
	\hline 
	& 1 & $1$ & $1$ & $1$\\
	& 2 & $4$ & $120$ & $2,036$\\
	& 3 & $\space$ & $1,191$ & $152,637$\\
	& 4 & $\space$ & $2,416$ & $2,203,488$\\
	& 5 & $\space$ & $\space$ & $9,738,114$\\
	& 6 & $\space$ & $\space$ & $15,724,248$\\
	\hline \hline
	\end{tabular}
 \caption {{\bf Cardinal B-Splines at the knot sequence}. Table entries are calculated using Eq.~\eqref{eq:MRA-recursiveBsplineatintegerpoints} and can be also found in \citet{Goswami1999}.} 
 \label{table:MRA-BsplineatIntegerPoints}
 \end{center}
 \end {table}

\subsection{Frequency Analysis}
To determine the level of approximation in the model, $j$, detailed in the previous section, that is the number of basis functions required to reconstruct the field, $v_t(r)$, the frequency response of the B-spline function needs to be computed by applying the Fourier transform to Eq.~\eqref{eq:N1convolutions}. To proceed the Fourier transform of $N_1(r)$ is first calculated, 
\begin{align}\label{eq:MRA-N1Fouriertransform}
\mathcal F(N_1(r))&=\int_{-\infty}^{+\infty}N_1(r)\mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\int_{0}^{1} \mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}.
\end{align}
The convolution theorem states that the Fourier transform of the convolution of functions is equal to the product of their Fourier transforms. Therefore, taking Fourier transform of Eq.~\eqref{eq:N1convolutions} and substituting Eq.~\eqref{eq:MRA-N1Fouriertransform} for $\mathcal F(N_1(r)) $ yields
\begin{align}\label{eq:MRA-NmFouriertransform}
\mathcal F(N_m(r))=\left(\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}\right)^m.
\end{align}
\dean{Need to say how the FT is used somewhere here before proceeding.} \parham{The Fourier transform of the $m$-th order B-spline allows the calculation of the spatial frequency response of the wavelets using Eq.~\eqref{eq:MRA-TwoScalepair2} giving
\begin{align}      
	  \mathcal{F}(\varphi_{m}\left(r\right)) &= \sum_{n=0}^{3m-2} \frac{\left(-1\right)^n}{2^{m-1}} \left[\mathbf q\right]_n \mathcal{F}\left(N_{m}\left(2r-n\right)\right) \nonumber\\
	&=\sum_{n=0}^{3m-2} \frac{\left(-1\right)^n}{2^{m}} \left[\mathbf q\right]_n \mathrm{exp}(-2\pi in\nu)\left(\frac{1-\mathrm{exp}(-\pi i \nu)}{\pi i\nu}\right)^m.\label{eq:MRA-Wavelettransform}
\end{align}
 From Eq.~\eqref{eq:MRA-Wavelettransform} the level of approximation, $j$, can be determined. This way we choose wavelets upto level $j$ which fully describe the spatial frequency contents of the underlying dynamics estimated from the observed field.}
